Efficient Machine Learning in a Memristor Network

### 2025-06-02 paper/paper.tex
Uploaded data for plot reproducing local

Modern computing algortihms are able to efficiently perform complicated nuemrical and symbolic operations, but their sequential structure limits their performance on more perceptual tasks. Inspired by the architecture and the functioning of the human brain that performs tasks via distributed and parallel fashion, artificial neural network (ANN) algorithms outreached traditional computing achieving remarkable success in challenging tasks ranging from image recognition to natural language processing and became central to many technological advancements in the field of artificial intelligence.

Recent advances in AI systems have been powered by the silicon-based computing power of the Graphics Processing Units (GPUs) and by an unprecedented abundance of data. With a clear trend of larger ANN models, their the energy requirements have grown significantly, negatively impacting their efficiency and scalability\cite{energy_cost_ml1, energy_cost_ml2}. With the rapid increase in computation requirements of AI training and the slower improvemnts in computing hardware traditionally predicted by Moore's Law, the necessity of new computing paradigms became more relevant.  As an alternative to ANN performed strictly in-silico, physical neural networks (PNN) have been introduced, where physical processes rather than mathematical operations are trained\cite{trainingphysicalneuralnetworks}. Since physical processes can be faster and more energetically efficient than their digital representation, this direction constitutes a promosing research direction to address this issue.

Applications of PNNs have been explored across diverse fields, including optics \cite{optics1}, electronics \cite{electronics1}, and fluidic and mechanical systems \cite{}, among others. Many existing training algorithms rely entirely on in silico methods, which often face challenges when transferring the learned model to real physical devices. To address this, hybrid training approaches have been proposed, where the inference phase is executed physically while the backward phase is computed numerically. Other methods, based on coupled learning, implement local learning rules at the level of each connection or edge. However, these approaches still face the need for two identical physical systems to apply different boundary conditions. In this work, we propose a novel algorithm in which both inference and backpropagation are carried out through physical processes, specifically leveraging the natural relaxation dynamics of the system in response to external stimuli and weight modifications.

A key advantage of PNNs is their ability to more closely emulate certain functional aspects of biological neural systems. One of the defining features of the brain is its capacity to process and integrate multiple types of stimuli—such as electrical, chemical, and mechanical inputs—simultaneously. This multimodal responsiveness is believed to contribute significantly to the brain’s efficiency and adaptability \cite{}. Inspired by this, we focus on neural networks constructed from memristive elements. Specifically, we utilize ionic microfluidic channels in aqueous solutions\cite{Rene_IonicNeuromorphic}, which exhibit the unique ability to respond concurrently to diverse physical stimuli. This characteristic is central to the operation of our proposed training algorithm.

@@

The paper is organized as follows. Section 2 reviews related work on memristor networks, machine learning algorithms, and physical learning methods. Section 3 describes the methods used in our study, including the network model, machine learning algorithm, and the physical learning approach. In Section 4, we present our proposed method for coupling memristors with capacitors and discuss the necessary properties for effective training. Section 5 details our experimental setup and implementation. The results and analysis are provided in Section 6, followed by conclusions and suggestions for future work in Section 7.

@@

\subsection{Training a Memristor Voltage Divider}

@@

\subsection{Training a memristor network}

@@

% \cleardoublepage
\subsection{Linear Regression}

>>>>

Modern computing algortihms are able to efficiently perform complicated nuemrical and symbolic operations, but their sequential structure limits their performance on more perceptual tasks. Inspired by the architecture and the functioning of the human brain, that performs tasks via distributed and parallel fashion, artificial neural network (ANN) algorithms outreached traditional computing achieving remarkable success in challenging tasks, such as visual and audio pattern recognition~\cite{pattern_recognition,review_image_rec, speech_rec} to natural language processing and became central to many technological advancements in the field of artificial intelligence\cite{ai_tech1}.

Recent developments in artificial intelligence have been driven by the computational power of Graphics Processing Units and the availability of large amounts of data. However, as ANNs continue to grow in size and complexity, their energy demands have increased substantially, ultimately impacting their efficiency and limiting their scalability \cite{energy_cost_ml1, energy_cost_ml2}. Meanwhile, the gap between the accelerating computational needs of AI training and the gradual improvements in traditional silicon-based hardware, as historically described by Moore’s Law, has highlighted the need for innovative computing architectures. One promising alternative is the development of physical neural networks (PNNs), which train physical processes rather than relying solely on digital calculations \cite{trainingphysicalneuralnetworks,backporp_pnn}. Because physical systems can often achieve faster and more energy-efficient operations than their digital representations, this direction constitutes a promising research direction to address this issue.

Applications of PNNs have been explored across diverse fields, including optics \cite{optics1}, electronics \cite{electronics1}, photonics \cite{acoustic1} and mechanical systems \cite{mechanics1}, among others. Many existing training algorithms rely entirely on in silico methods, which often face challenges when transferring the learned model to real physical devices. To address this, hybrid training approaches have been proposed, where the inference phase is executed physically while the backward phase is computed numerically\cite{backporp_pnn}. Other methods, based on coupled learning\cite{supervised_learning}, implement local learning rules at the level of each connection or edge\cite{demonstr_coupled_learning}. Applying local rules removed the need of the optimiziation of a global function during training, as in standard backpropagation methods. However, these approaches still face the need for two identical physical systems to apply different boundary conditions. In this work, we propose a novel algorithm in which both inference and backpropagation are carried out mainly taking advantage of the physical processes inherent in the system, specifically leveraging the natural relaxation dynamics of the system in response to external stimuli and weight modifications.

A key advantage of PNNs is their ability to more closely emulate certain functional aspects of biological neural systems. One of the defining features of the brain is its capacity to process and integrate multiple types of stimuli, such as electrical, chemical, and mechanical inputs, simultaneously. This multimodal responsiveness is believed to contribute significantly to the brain’s efficiency and adaptability \cite{}. Inspired by this functionality, we base our approach on physical neural networks constructed from memristive elements. Specifically, we utilize ionic microfluidic channels in aqueous solutions\cite{Rene_IonicNeuromorphic}, which exhibit the ability to respond concurrently to diverse physical stimuli. This characteristic is central to the operation of our proposed training algorithm.

@@

The paper is organized as follows. \cref{sec:memristor_network} presents the physical network we define using memristor elements, called memristor network, and explains how this system is numerically solved. \cref{sec:training} describes the training algorithm that we use to achieve desired responses of the memristor network. In \cref{sec:train_memr_vd}, we exemplify the proposed algorithm by training a simple system, namely a voltage divider. \cref{sec:train_memr_nw} s The results and analysis are provided in Section 6, followed by conclusions and suggestions for future work in Section 7.

@@

\subsection{\label{sec:train_memr_vd}Training a Memristor Voltage Divider}

@@

\subsection{\label{sec:train_memr_nw}Training a memristor network}

@@

% \cleardoublepage
\subsection{\label{sec:linear_regression}Linear Regression}
